% ══════════════════════════════════════════════════════════════════
% Byzantine-Based Blockchain Consensus for EV-to-Grid Energy Trading:
% A Static-Temporal Security Evaluation Framework
% ══════════════════════════════════════════════════════════════════
\documentclass[10pt,twocolumn,journal]{IEEEtran}

% ─── Packages ────────────────────────────────────────────────────
\usepackage{amsmath,amssymb,amsthm,mathtools}
\usepackage{bm,dsfont}
\usepackage{booktabs,multirow,array,colortbl,tabularx}
\usepackage{graphicx,subcaption}
\graphicspath{{figures/}{premium/}}
\usepackage{geometry}
\geometry{margin=0.75in}
\usepackage{xcolor,enumitem}
\definecolor{ieee}{RGB}{0,83,159}
\usepackage[colorlinks=true, allcolors=ieee]{hyperref}
\usepackage{cite}
\usepackage{url}
\usepackage{tikz}
\usetikzlibrary{shapes.geometric, arrows, positioning, fit, backgrounds, calc}
\usepackage{flushend}

\newtheorem{theorem}{Theorem}
\newtheorem{definition}{Definition}
\newtheorem{proposition}{Proposition}
\newtheorem{lemma}{Lemma}

\begin{document}

\title{\Huge\bfseries Byzantine-Based Blockchain Consensus for EV-to-Grid Energy Trading:\\A Static-Temporal Security Evaluation Framework}

\author{\textbf{Atharv~Manojkumar~Shukla}\\
\vspace{0.15cm}
\small \emph{Co-Authors / Project Guides:} \textbf{Prof.~Uday~Suryavanshi} and \textbf{Dr.~Sunny~Kumar}%
\thanks{A. M. Shukla is a 3rd year B.Tech Undergraduate Student with the Department of Computer Science and Engineering, Indian Institute of Information Technology (IIIT), Nagpur 441108, India (e-mail: atharv@iiitn.ac.in).}%
\thanks{Prof. Uday Suryavanshi and Dr. Sunny Kumar are Co-Authors \& Project Guides with the E-MC$^2$ Laboratory, Department of Electrical Engineering, Veermata Jijabai Technological Institute (VJTI), Mumbai 400019, India (e-mail: usuryavanshi@ee.vjti.ac.in, skumar@ee.vjti.ac.in).}}

\markboth{Research Project Report --- E-MC$^2$ Laboratory, VJTI Mumbai}%
{Shukla \MakeLowercase{\textit{et al.}}: Byzantine-Based Blockchain Consensus for EV-to-Grid Energy Trading}

\maketitle

% ═══════════════════════════════════════════════════════════════
% ABSTRACT
% ═══════════════════════════════════════════════════════════════
\begin{abstract}
The rapid integration of Electric Vehicles (EVs) into smart Advanced Metering Infrastructure (AMI) enables decentralized peer-to-peer (P2P) energy trading, but exposes power distribution networks to severe cyber-physical vulnerabilities. Moving beyond simple consensus benchmarking, this paper presents a comprehensive engineering research paper structured around the \emph{Static–Temporal Security Framework (STSF)}. We address the fundamental research question: \emph{How does introducing Byzantine consensus transform the security of an EV energy trading system compared to its baseline centralized architecture, and how do consensus mechanisms alter this security-latency trade-off?} Using an IEEE 33-bus distribution system with 50 EVs and 3,854 physical sensors as an operational reference, we show that under baseline model assumptions, permissioned Byzantine Fault Tolerant (BFT) blockchain consensus reduces total attack compromise probability from $P_{\text{TA}} \approx 0.005$ to $P_{\text{TAb}} \approx 10^{-173}$ (a 170 order-of-magnitude static gain under the Sheikh formulation) and lowers single-point-of-failure risk ($P_{\text{SPOF}} \to 10^{-10}$). We expand defensive coverage across a 12-attack cyber-physical vector taxonomy and model how consensus validation latency induces a Poisson temporal vulnerability decay ($W(\tau)$). Sub-second pipelined consensus protocols (Tower BFT, RVR) preserve over 92\% of static security gains ($P_{\text{secure}} \ge 0.927$), whereas classical multi-round protocols ($OM(m)$) collapse due to latency exposure ($P_{\text{secure}} = 0.012$). $10^6$ Monte Carlo simulation trials corroborate analytical derivations to within 0.5\% error with 95\% confidence intervals, providing an actionable consensus selection roadmap for smart grid practitioners.
\end{abstract}

\begin{IEEEkeywords}
Advanced Metering Infrastructure (AMI), Electric Vehicle (EV) Energy Trading, Byzantine Fault Tolerance (BFT), Static–Temporal Security Framework (STSF), Cyber-Physical Systems, Poisson Vulnerability Decay.
\end{IEEEkeywords}

% ═══════════════════════════════════════════════════════════════
\section{Introduction}
\label{sec:introduction}
% ═══════════════════════════════════════════════════════════════

\IEEEPARstart{T}{he} global transition toward smart grid decentralization has positioned Electric Vehicles (EVs) not merely as energy consumers, but as active prosumers capable of providing Vehicle-to-Grid (V2G) power balancing and ancillary services. Advanced Metering Infrastructure (AMI) acts as the primary digital backbone enabling high-frequency telemetry and peer-to-peer (P2P) energy settlement. Fig. \ref{fig:fig1_evolution} illustrates the historical evolution of EV-to-grid energy trading architectures from traditional one-way distribution networks to cyber-physical smart grid markets.

\begin{figure}[htbp]
\centering
\resizebox{\columnwidth}{!}{%
\begin{tikzpicture}[node distance=0.7cm, auto, >=latex]
    \node [draw, rectangle, fill=blue!10, text width=6.5cm, text centered, rounded corners] (g1) {\textbf{1. Traditional Grid} \\ One-way Power Flow, Centralized Generation};
    \node [draw, rectangle, fill=blue!20, text width=6.5cm, text centered, rounded corners, below=0.3cm of g1] (g2) {\textbf{2. Smart Grid} \\ Bidirectional Telemetry, Automated Meter Reading};
    \node [draw, rectangle, fill=blue!30, text width=6.5cm, text centered, rounded corners, below=0.3cm of g2] (g3) {\textbf{3. Advanced Metering Infrastructure (AMI)} \\ Real-time Demand Response, High-Frequency Telemetry};
    \node [draw, rectangle, fill=orange!20, text width=6.5cm, text centered, rounded corners, below=0.3cm of g3] (g4) {\textbf{4. EV Integration (V2G)} \\ Mobile Energy Storage, Distributed Prosumer Trading};
    \node [draw, rectangle, fill=green!20, text width=6.5cm, text centered, rounded corners, below=0.3cm of g4] (g5) {\textbf{5. P2P Energy Trading} \\ Direct EV-to-Grid Settlement, Dynamic Tariffs};
    \node [draw, rectangle, fill=green!35, text width=6.5cm, text centered, rounded corners, below=0.3cm of g5] (g6) {\textbf{6. Cyber-Physical Energy Trading (STSF)} \\ Proactive BFT Validation Layer, Fault-Tolerant Grid};

    \draw [->, thick] (g1) -- (g2);
    \draw [->, thick] (g2) -- (g3);
    \draw [->, thick] (g3) -- (g4);
    \draw [->, thick] (g4) -- (g5);
    \draw [->, thick] (g5) -- (g6);
\end{tikzpicture}%
}
\caption{\textbf{Figure 1 --- Evolution of EV-to-Grid Energy Trading:} Illustrating the operational transition from traditional one-way power distribution to cyber-physical BFT-secured P2P energy trading.}
\label{fig:fig1_evolution}
\end{figure}

\subsection{Engineering Motivation & Operational Workflow}
In a baseline EV energy trading system without dedicated cybersecurity validation, transactions follow the operational workflow depicted in Fig. \ref{fig:fig2_workflow_no_cyber}. Smart meters log charging rates, forward telemetry across cellular/mesh links to a central Meter Data Management System (MDMS) coordinator, which updates SCADA databases and executes grid dispatch commands.

\begin{figure}[htbp]
\centering
\resizebox{\columnwidth}{!}{%
\begin{tikzpicture}[node distance=0.6cm, auto, >=latex]
    \node [draw, rectangle, fill=blue!15, text width=1.8cm, text centered, rounded corners] (ev) {\textbf{1. EV Prosumer}};
    \node [draw, rectangle, fill=blue!25, text width=1.8cm, text centered, rounded corners, right=0.4cm of ev] (sm) {\textbf{2. Smart Meter}};
    \node [draw, rectangle, fill=orange!20, text width=1.8cm, text centered, rounded corners, right=0.4cm of sm] (comm) {\textbf{3. Network Comm.}};
    \node [draw, rectangle, fill=red!20, text width=1.8cm, text centered, rounded corners, right=0.4cm of comm] (coord) {\textbf{4. Central Coordinator}};

    \node [draw, rectangle, fill=red!30, text width=4.0cm, text centered, rounded corners, below=0.5cm of coord] (scada) {\textbf{5. Central SCADA / MDMS}};
    \node [draw, rectangle, fill=green!20, text width=4.0cm, text centered, rounded corners, left=0.6cm of scada] (grid) {\textbf{6. Physical Grid Control}};

    \draw [->, thick] (ev) -- (sm);
    \draw [->, thick] (sm) -- (comm);
    \draw [->, thick] (comm) -- (coord);
    \draw [->, thick] (coord) -- (scada);
    \draw [->, thick] (scada) -- (grid);

    % Callout boxes highlighting limitations
    \node [draw, rectangle, dashed, red!80!black, fill=red!5, text width=8.0cm, text centered, below=0.4cm of scada] (vun) {\textbf{Critical Engineering Vulnerabilities (Without Cybersecurity):} \\ (1) Absolute SPOF ($P_{\text{SPOF}}=1.0$), (2) Zero Byzantine Tolerance ($f=0$), \\ (3) Unverified Data Injection, (4) Central Key File Leakage Risk};
\end{tikzpicture}%
}
\caption{\textbf{Figure 2 --- Engineering Workflow (Without Cybersecurity):} Operational data flow in centralized energy trading highlighting single points of failure and coordinator vulnerabilities.}
\label{fig:fig2_workflow_no_cyber}
\end{figure}

As shown in Fig. \ref{fig:fig2_workflow_no_cyber}, this centralized engineering workflow suffers from fundamental structural vulnerabilities:
\begin{enumerate}[leftmargin=*]
    \item \textbf{Single Point of Failure (SPOF):} The central coordinator host presents a single point of failure ($P_{\text{SPOF}} = 1.0$). If breached by an adversary, the entire billing and control perimeter collapses.
    \item \textbf{Zero Byzantine Resilience:} Centralized databases cannot detect or reject corrupt commands if the central coordinator node itself becomes Byzantine or compromised ($f=0$).
    \item \textbf{Unvalidated Sensor Ingestion:} Measurement readings are trusted upon arrival, enabling False Data Injection (FDI) to alter state estimation.
\end{enumerate}

\subsection{Research Question & Core Objectives}
To address these limitations, this paper investigates the central research question:

\begin{quote}
\textbf{``How does introducing Byzantine consensus transform the engineering characteristics of the original EV-to-grid energy trading framework, and how can this improvement be analytically quantified across different consensus mechanisms?''}
\end{quote}

Rather than presenting consensus benchmarking as an isolated exercise, we unify engineering system modeling, threat analysis, analytical probability derivation, and temporal latency exposure into the \emph{Static–Temporal Security Framework (STSF)}.

% ═══════════════════════════════════════════════════════════════
\section{Cyber-Physical Attack Surface & Taxonomy}
\label{sec:threat_model}
% ═══════════════════════════════════════════════════════════════

Fig. \ref{fig:fig3_attack_surface} illustrates the comprehensive 12-attack cyber-physical attack surface categorized by system layer: Physical Layer, Communication Layer, Application Layer, and Consensus Layer.

\begin{figure}[htbp]
\centering
\resizebox{\columnwidth}{!}{%
\begin{tikzpicture}[node distance=0.5cm, auto, >=latex]
    % Physical Layer
    \node [draw, rectangle, fill=red!15, text width=8.2cm, rounded corners] (phys) {\textbf{1. Physical Layer Attacks} \\ $\bullet$ Sensor Compromise ($P_{\text{SA}}$) \quad $\bullet$ Receiver/Actuator Override ($P_{\text{R}}$)};
    
    % Comm Layer
    \node [draw, rectangle, fill=orange!15, text width=8.2cm, rounded corners, below=0.3cm of phys] (comm) {\textbf{2. Communication Layer Attacks} \\ $\bullet$ Comm. Hijack ($P_{\text{CA}}$) \quad $\bullet$ Man-in-the-Middle ($P_{\text{MitM}}$) \quad $\bullet$ Replay Vector ($P_{\text{Replay}}$) \\ $\bullet$ Port DoS ($P_{\text{DoS}}$) \quad $\bullet$ Botnet DDoS ($P_{\text{DDoS}}$)};
    
    % App Layer
    \node [draw, rectangle, fill=yellow!20, text width=8.2cm, rounded corners, below=0.3cm of comm] (app) {\textbf{3. Application & Control Layer Attacks} \\ $\bullet$ SCADA Breach ($P_{\text{SCADA}}$) \quad $\bullet$ Private Key Compromise ($P_{\text{Key}}$) \quad $\bullet$ False Data Injection ($P_{\text{FDI}}$)};
    
    % Consensus Layer
    \node [draw, rectangle, fill=green!20, text width=8.2cm, rounded corners, below=0.3cm of app] (cons) {\textbf{4. Consensus & Identity Layer Attacks} \\ $\bullet$ Sybil Identity Spoofing ($P_{\text{Sybil}}$) \quad $\bullet$ Byzantine Node Collusion ($P_{\text{Byz}}$)};

    \draw [->, thick] (phys) -- (comm);
    \draw [->, thick] (comm) -- (app);
    \draw [->, thick] (app) -- (cons);
\end{tikzpicture}%
}
\caption{\textbf{Figure 3 --- Cyber Attack Surface Taxonomy:} Structuring 12 cyber-physical attack vectors across Physical, Communication, Application, and Consensus layers.}
\label{fig:fig3_attack_surface}
\end{figure}

Fig. \ref{fig:fig4_eng_vs_cyber} contrasts the original engineering system against the proposed cybersecurity-enhanced architecture.

\begin{figure}[htbp]
\centering
\resizebox{\columnwidth}{!}{%
\begin{tikzpicture}[node distance=0.6cm, auto, >=latex]
    \node [draw, rectangle, fill=blue!10, text width=6.5cm, text centered, rounded corners] (e1) {\textbf{Original Engineering System} \\ Centralized Metering \& SCADA Dispatch};
    \node [draw, rectangle, fill=red!10, text width=6.5cm, text centered, rounded corners, below=0.3cm of e1] (e2) {\textbf{Engineering Limitations} \\ Single Point of Failure ($P_{\text{SPOF}}=1.0$), Zero Byzantine Tolerance};
    \node [draw, rectangle, fill=yellow!20, text width=6.5cm, text centered, rounded corners, below=0.3cm of e2] (e3) {\textbf{Cybersecurity Layer (STSF)} \\ Proactive Validation \& Threshold Cryptography};
    \node [draw, rectangle, fill=green!15, text width=6.5cm, text centered, rounded corners, below=0.3cm of e3] (e4) {\textbf{Blockchain Consensus Engine} \\ Distributed Validator Quorums ($n=51, f=16$)};
    \node [draw, rectangle, fill=green!30, text width=6.5cm, text centered, rounded corners, below=0.3cm of e4] (e5) {\textbf{Improved Trust & Grid Reliability} \\ 170 Orders-of-Magnitude Gain, $P_{\text{SPOF}} \to 10^{-10}$};

    \draw [->, thick] (e1) -- (e2);
    \draw [->, thick] (e2) -- (e3);
    \draw [->, thick] (e3) -- (e4);
    \draw [->, thick] (e4) -- (e5);
\end{tikzpicture}%
}
\caption{\textbf{Figure 4 --- Engineering vs. Cybersecurity:} Architectural transformation showing how integrating a BFT validation layer resolves centralized engineering limitations.}
\label{fig:fig4_eng_vs_cyber}
\end{figure}

% ═══════════════════════════════════════════════════════════════
\section{Analytical Security Framework (STSF)}
\label{sec:framework_stsf}
% ═══════════════════════════════════════════════════════════════

Fig. \ref{fig:fig5_research_framework} outlines the overall research framework connecting engineering modeling, threat taxonomy, static reliability, temporal exposure, and consensus selection.

\begin{figure}[htbp]
\centering
\resizebox{\columnwidth}{!}{%
\begin{tikzpicture}[node distance=0.5cm, auto, >=latex]
    \node [draw, rectangle, fill=blue!15, text width=8.0cm, text centered, rounded corners] (f1) {\textbf{1. Engineering System Model} (IEEE 33-Bus + 50 EVs + 3854 Sensors)};
    \node [draw, rectangle, fill=red!15, text width=8.0cm, text centered, rounded corners, below=0.3cm of f1] (f2) {\textbf{2. 12-Attack Cyber-Physical Threat Model}};
    \node [draw, rectangle, fill=orange!15, text width=8.0cm, text centered, rounded corners, below=0.3cm of f2] (f3) {\textbf{3. Analytical STSF Formulation} ($P_{\text{secure}} = f(P_{\text{TAb}}, \tau, \lambda)$)};
    \node [draw, rectangle, fill=yellow!20, text width=8.0cm, text centered, rounded corners, below=0.3cm of f3] (f4) {\textbf{4. Static Reliability Analysis} (170 Orders-of-Magnitude Gain)};
    \node [draw, rectangle, fill=green!15, text width=8.0cm, text centered, rounded corners, below=0.3cm of f4] (f5) {\textbf{5. Temporal Poisson Exposure Analysis} ($W(\tau) = 1 - e^{-\lambda \tau P_{\text{TA}}}$)};
    \node [draw, rectangle, fill=green!25, text width=8.0cm, text centered, rounded corners, below=0.3cm of f5] (f6) {\textbf{6. Consensus Benchmark & Trade-off Matrix} (9 Protocols)};
    \node [draw, rectangle, fill=green!40, text width=8.0cm, text centered, rounded corners, below=0.3cm of f6] (f7) {\textbf{7. Practitioner Deployment Roadmap}};

    \draw [->, thick] (f1) -- (f2);
    \draw [->, thick] (f2) -- (f3);
    \draw [->, thick] (f3) -- (f4);
    \draw [->, thick] (f4) -- (f5);
    \draw [->, thick] (f5) -- (f6);
    \draw [->, thick] (f6) -- (f7);
\end{tikzpicture}%
}
\caption{\textbf{Figure 5 --- Overall Research Framework:} Complete analytical roadmap from engineering grid modeling to practical deployment recommendation.}
\label{fig:fig5_research_framework}
\end{figure}

Fig. \ref{fig:fig6_methodology} details the step-by-step methodology pipeline governing mathematical modeling, validation, sensitivity sweeps, and Monte Carlo verification.

\begin{figure}[htbp]
\centering
\resizebox{\columnwidth}{!}{%
\begin{tikzpicture}[node distance=0.5cm, auto, >=latex]
    \node [draw, rectangle, fill=blue!10, text width=8.0cm, rounded corners] (m1) {\textbf{1. Engineering & Threat Assumptions} ($x=0.95, m=10, p_c=0.05$)};
    \node [draw, rectangle, fill=orange!15, text width=8.0cm, rounded corners, below=0.3cm of m1] (m2) {\textbf{2. Closed-Form Probability Derivation} (Binomial Quorums \& $M/M/1$)};
    \node [draw, rectangle, fill=yellow!20, text width=8.0cm, rounded corners, below=0.3cm of m2] (m3) {\textbf{3. Parameter Sensitivity Sweeps} ($x \in [0.90, 0.999], \lambda \in [1, 100]$)};
    \node [draw, rectangle, fill=green!20, text width=8.0cm, rounded corners, below=0.3cm of m3] (m4) {\textbf{4. Statistically Rigorous Monte Carlo Verification} ($10^6$ Trials)};
    \node [draw, rectangle, fill=green!35, text width=8.0cm, rounded corners, below=0.3cm of m4] (m5) {\textbf{5. Scholarly Synthesis & Engineering Principles}};

    \draw [->, thick] (m1) -- (m2);
    \draw [->, thick] (m2) -- (m3);
    \draw [->, thick] (m3) -- (m4);
    \draw [->, thick] (m4) -- (m5);
\end{tikzpicture}%
}
\caption{\textbf{Figure 6 --- Methodology Pipeline:} Methodological progression from mathematical assumptions to Monte Carlo verification.}
\label{fig:fig6_methodology}
\end{figure}

Fig. \ref{fig:fig7_without_vs_with} presents a side-by-side architectural comparison contrasting the centralized vulnerability structure against the distributed BFT consensus validation engine.

\begin{figure*}[t]
\centering
\resizebox{\textwidth}{!}{%
\begin{tikzpicture}[node distance=0.8cm, auto, >=latex]
    % Panel A: Without Blockchain
    \node [draw, rectangle, fill=red!10, text width=7.5cm, minimum height=4.5cm, rounded corners] (panelA) {};
    \node [font=\bfseries, color=red!80!black, anchor=north west] at (panelA.north west) {\quad Panel A: Centralized Architecture (Without Blockchain)};
    \node [draw, rectangle, fill=red!25, text width=6.5cm, text centered, rounded corners, yshift=0.5cm] at (panelA.center) (c_coord) {\textbf{Central Coordinator / MDMS} \\ Single Point of Failure ($P_{\text{SPOF}}=1.0$) \\ Unverified Telemetry Ingestion};
    \node [draw, rectangle, fill=red!40, text width=6.5cm, text centered, rounded corners, below=0.5cm of c_coord] (c_fail) {\textbf{Attack Compromise Success} \\ $P_{\text{Compromise}} = 0.974$ \quad ($P_{\text{Sybil}}=1.0, P_{\text{Byz}}=1.0$)};
    \draw [->, ultra thick, red!80!black] (c_coord) -- (c_fail);

    % Panel B: With Blockchain
    \node [draw, rectangle, fill=green!10, text width=9.0cm, minimum height=4.5cm, rounded corners, right=1.0cm of panelA] (panelB) {};
    \node [font=\bfseries, color=green!60!black, anchor=north west] at (panelB.north west) {\quad Panel B: Distributed BFT Architecture (With Blockchain)};
    \node [draw, rectangle, fill=green!25, text width=8.0cm, text centered, rounded corners, yshift=0.5cm] at (panelB.center) (b_nodes) {\textbf{Distributed Validator Network ($n=51, f=16$)} \\ $2f+1$ Multi-Signature Voting Quorums \\ Threshold Secret Sharing \& PoH Timestamps};
    \node [draw, rectangle, fill=green!40, text width=8.0cm, text centered, rounded corners, below=0.5cm of b_nodes] (b_succ) {\textbf{Proactive Attack Rejection} \\ $P_{\text{TAb}} \approx 10^{-173}$ (170 Orders Gain), $P_{\text{SPOF}} \to 10^{-10}$};
    \draw [->, ultra thick, green!60!black] (b_nodes) -- (b_succ);
\end{tikzpicture}%
}
\caption{\textbf{Figure 7 --- Without Blockchain vs. With Blockchain Architecture:} Comparative architectural representation contrasting centralized coordinator failure with distributed BFT voting quorums.}
\label{fig:fig7_without_vs_with}
\end{figure*}

Fig. \ref{fig:fig8_prob_modeling} illustrates the probability modeling pipeline connecting physical engineering causes to variables, assumptions, equations, and overall security metrics.

\begin{figure}[htbp]
\centering
\resizebox{\columnwidth}{!}{%
\begin{tikzpicture}[node distance=0.5cm, auto, >=latex]
    \node [draw, rectangle, fill=blue!15, text width=8.0cm, rounded corners] (p1) {\textbf{1. Physical Engineering Cause} (Physical Meter Tampering / Network Sniffing)};
    \node [draw, rectangle, fill=orange!15, text width=8.0cm, rounded corners, below=0.3cm of p1] (p2) {\textbf{2. Model Variables} (Sensor Accuracy $x$, Local Subset $m$, Fault Bound $f$)};
    \node [draw, rectangle, fill=yellow!20, text width=8.0cm, rounded corners, below=0.3cm of p2] (p3) {\textbf{3. Operational Assumptions} (Independent Compromise Rate $p_c=0.05$)};
    \node [draw, rectangle, fill=green!15, text width=8.0cm, rounded corners, below=0.3cm of p3] (p4) {\textbf{4. Closed-Form Derivation} ($P_{\text{Byz}} = \sum_{k=f+1}^n \binom{n}{k} p_c^k (1-p_c)^{n-k}$)};
    \node [draw, rectangle, fill=green!35, text width=8.0cm, rounded corners, below=0.3cm of p4] (p5) {\textbf{5. Security Metric Output} ($P_{\text{secure}}(\lambda, \tau) = (1-P_{\text{TAb}}) e^{-\lambda \tau P_{\text{TA}}}$)};

    \draw [->, thick] (p1) -- (p2);
    \draw [->, thick] (p2) -- (p3);
    \draw [->, thick] (p3) -- (p4);
    \draw [->, thick] (p4) -- (p5);
\end{tikzpicture}%
}
\caption{\textbf{Figure 8 --- Probability Modeling Pipeline:} Systematic progression showing how physical causes map into mathematical equations and security metrics.}
\label{fig:fig8_prob_modeling}
\end{figure}

\subsection{Closed-Form Complexity-to-Latency Derivation $\tau = f(M, n)$}
To map protocol message complexity into physical network queueing, we derive validation latency $\tau(n)$ combining propagation delay, message complexity transmission delay, and $M/M/1$ queueing delay (Kleinrock 1975):

\begin{equation}
\tau(n) = \tau_{\text{prop}} + \frac{M(n) \cdot S_{\text{msg}}}{B_{\text{net}}} + \frac{\lambda \cdot S_{\text{msg}}}{\mu - \lambda \cdot S_{\text{msg}}}
\label{eq:latency_derivation_restructured}
\end{equation}

where $M(n)$ represents total consensus message exchanges. We note that Lamport's $OM(m)$ algorithm ($\tau = 43.35\text{s}$) serves solely as the theoretical origin of Byzantine agreement and is included to illustrate historical consensus evolution rather than as a practical deployment candidate.

% ═══════════════════════════════════════════════════════════════
\section{Consensus Selection Roadmap & Trade-offs}
\label{sec:consensus_selection}
% ═══════════════════════════════════════════════════════════════

Fig. \ref{fig:fig9_decision_tree} presents the practical consensus selection decision tree guiding utility engineers based on network latency, security requirements, and validator scale.

\begin{figure}[htbp]
\centering
\resizebox{\columnwidth}{!}{%
\begin{tikzpicture}[node distance=0.8cm, auto, >=latex]
    \node [draw, diamond, fill=blue!15, text width=2.5cm, text centered] (q1) {\textbf{Sub-Second Latency?}};
    \node [draw, diamond, fill=orange!15, text width=2.5cm, text centered, below left=0.8cm and 0.3cm of q1] (q2) {\textbf{High Scalability ($n>50$)?}};
    \node [draw, diamond, fill=yellow!20, text width=2.5cm, text centered, below right=0.8cm and 0.3cm of q1] (q3) {\textbf{Clustered Grid Topology?}};

    \node [draw, rectangle, fill=green!35, text width=2.8cm, text centered, rounded corners, below=0.6cm of q2] (ans_g4) {\textbf{Select G4 Protocol} \\ Tower BFT / RVR ($\mathcal{O}(n)$ Linear)};
    \node [draw, rectangle, fill=green!20, text width=2.8cm, text centered, rounded corners, below=0.6cm of q3] (ans_g3) {\textbf{Select G3 Protocol} \\ G-PBFT / SV-PBFT ($\mathcal{O}(n_c^2)$ Clustered)};

    \draw [->, thick] (q1) -- node [left, font=\scriptsize] {Yes} (q2);
    \draw [->, thick] (q1) -- node [right, font=\scriptsize] {No} (q3);
    \draw [->, thick] (q2) -- node [left, font=\scriptsize] {Yes} (ans_g4);
    \draw [->, thick] (q3) -- node [right, font=\scriptsize] {Yes} (ans_g3);
\end{tikzpicture}%
}
\caption{\textbf{Figure 9 --- Consensus Protocol Selection Framework:} Engineering decision tree guiding optimal protocol selection for smart grid AMI deployments.}
\label{fig:fig9_decision_tree}
\end{figure}

Fig. \ref{fig:fig10_static_vs_temporal} illustrates the transition from static security gains to temporal survival under consensus exposure latency.

\begin{figure}[htbp]
\centering
\resizebox{\columnwidth}{!}{%
\begin{tikzpicture}[node distance=0.5cm, auto, >=latex]
    \node [draw, rectangle, fill=blue!15, text width=8.0cm, rounded corners] (s1) {\textbf{1. Static Security Gain} ($P_{\text{TAb}} \approx 10^{-173}$, 170 Orders Gain)};
    \node [draw, rectangle, fill=orange!15, text width=8.0cm, rounded corners, below=0.3cm of s1] (s2) {\textbf{2. Consensus Validation Latency} ($\tau(n) = f(M, n)$)};
    \node [draw, rectangle, fill=red!15, text width=8.0cm, rounded corners, below=0.3cm of s2] (s3) {\textbf{3. Poisson Exposure Window} ($W(\tau) = 1 - e^{-\lambda \tau P_{\text{TA}}}$)};
    \node [draw, rectangle, fill=yellow!20, text width=8.0cm, rounded corners, below=0.3cm of s3] (s4) {\textbf{4. Temporal Vulnerability Risk} ($P_{\text{temporal}} \in [0.01, 0.98]$)};
    \node [draw, rectangle, fill=green!30, text width=8.0cm, rounded corners, below=0.3cm of s4] (s5) {\textbf{5. Realized System Survival} ($P_{\text{secure}} \ge 0.927$ for G4 Engines)};

    \draw [->, thick] (s1) -- (s2);
    \draw [->, thick] (s2) -- (s3);
    \draw [->, thick] (s3) -- (s4);
    \draw [->, thick] (s4) -- (s5);
\end{tikzpicture}%
}
\caption{\textbf{Figure 10 --- Static-to-Temporal Security Transition:} Conceptual bridge showing how static security gains decay under consensus validation latency.}
\label{fig:fig10_static_vs_temporal}
\end{figure}

Fig. \ref{fig:fig11_tradeoffs} highlights the multi-dimensional consensus engineering trade-offs balancing Latency, Throughput, Security, Communication Cost, and Scalability.

\begin{figure}[htbp]
\centering
\resizebox{\columnwidth}{!}{%
\begin{tikzpicture}[node distance=0.6cm, auto, >=latex]
    \node [draw, circle, fill=blue!20, minimum size=1.8cm, text centered] (sec) {\textbf{Security}};
    \node [draw, circle, fill=green!20, minimum size=1.8cm, text centered, right=1.2cm of sec] (lat) {\textbf{Latency}};
    \node [draw, circle, fill=orange!20, minimum size=1.8cm, text centered, below=1.0cm of sec] (tps) {\textbf{Throughput}};
    \node [draw, circle, fill=yellow!25, minimum size=1.8cm, text centered, right=1.2cm of tps] (cost) {\textbf{Comm. Cost}};

    \draw [<->, ultra thick, ieee] (sec) -- (lat);
    \draw [<->, ultra thick, ieee] (sec) -- (tps);
    \draw [<->, ultra thick, ieee] (lat) -- (cost);
    \draw [<->, ultra thick, ieee] (tps) -- (cost);
    \draw [<->, ultra thick, ieee] (sec) -- (cost);
    \draw [<->, ultra thick, ieee] (lat) -- (tps);
\end{tikzpicture}%
}
\caption{\textbf{Figure 11 --- Consensus Engineering Trade-offs:} Multi-dimensional trade-off matrix illustrating interdependencies between security, latency, throughput, and communication cost.}
\label{fig:fig11_tradeoffs}
\end{figure}

Fig. \ref{fig:fig12_deployment_roadmap} presents the practical utility deployment roadmap for implementing BFT consensus in production AMI networks.

\begin{figure}[htbp]
\centering
\resizebox{\columnwidth}{!}{%
\begin{tikzpicture}[node distance=0.5cm, auto, >=latex]
    \node [draw, rectangle, fill=blue!15, text width=8.0cm, rounded corners] (r1) {\textbf{Step 1: Engineering Assessment} (Map DCUs & Substation Gateways)};
    \node [draw, rectangle, fill=orange!15, text width=8.0cm, rounded corners, below=0.3cm of r1] (r2) {\textbf{Step 2: Threat Matrix Calibration} (Parameterize $x, y, z, p_c$)};
    \node [draw, rectangle, fill=yellow!20, text width=8.0cm, rounded corners, below=0.3cm of r2] (r3) {\textbf{Step 3: Select Consensus Engine} (Choose Tower BFT / RVR for $\mathcal{O}(n)$)};
    \node [draw, rectangle, fill=green!20, text width=8.0cm, rounded corners, below=0.3cm of r3] (r4) {\textbf{Step 4: Configure Validator Quorums} (Set $n=51, f=16$, HSM Keys)};
    \node [draw, rectangle, fill=green!35, text width=8.0cm, rounded corners, below=0.3cm of r4] (r5) {\textbf{Step 5: Production Deployment & Monitoring} (HIL 5G Testbed Verification)};

    \draw [->, thick] (r1) -- (r2);
    \draw [->, thick] (r2) -- (r3);
    \draw [->, thick] (r3) -- (r4);
    \draw [->, thick] (r4) -- (r5);
\end{tikzpicture}%
}
\caption{\textbf{Figure 12 --- Practical Deployment Roadmap:} Step-by-step guidance for utility practitioners deploying permissioned BFT consensus.}
\label{fig:fig12_deployment_roadmap}
\end{figure}

% ═══════════════════════════════════════════════════════════════
\section{Results & Hypothesis Verification}
\label{sec:results_restructured}
% ═══════════════════════════════════════════════════════════════

\subsection{Static Security Gains & SPOF Elimination}
Within the calibrated probabilistic model, applying permissioned BFT consensus onto AMI energy trading networks reduces Single Point of Failure risk from $P_{\text{SPOF}} = 1.0$ (standalone coordinator) to $P_{\text{Byz}} = 1.54 \times 10^{-10}$ under baseline parameters ($n=51, f=16, p_c=0.05$). Replay attacks are completely eliminated ($P_{\text{Replay}} \to 0$), SCADA breaches are reduced by 18 orders of magnitude, and key compromises are reduced by $1,015\times$. Under baseline probabilistic model assumptions ($N_{\text{sen}} = 3,854$), BFT consensus achieves a static security gain of 170 orders of magnitude ($P_{\text{TA}} \approx 0.005 \to P_{\text{TAb}} \approx 10^{-173}$).

\subsection{Temporal Vulnerability & Protocol Rankings}
As detailed in Table \ref{tab:protocol_ranking_restructured}, sub-second Generation-4 protocols (RVR: $P_{\text{secure}} = 0.9404, \tau = 200\text{ms}$; Tower BFT: $P_{\text{secure}} = 0.9272, \tau = 242.9\text{ms}$) preserve over 92\% of static security gains by restricting temporal vulnerability decay ($P_{\text{temporal}} < 0.024$). Conversely, high-latency Generation-1 protocols (Classic PBFT: $\tau = 7.65\text{s}$) suffer severe security decay ($P_{\text{secure}} = 0.4421$).

\begin{table*}[t]
\caption{Consensus Protocol Temporal Security Ranking for AMI Energy Trading ($\lambda = 20\text{ attacks/s}, n=51, f=16$)}
\label{tab:protocol_ranking_restructured}
\centering
\scriptsize
\begin{tabularx}{\textwidth}{@{}clrrrrcX@{}}
\toprule
\textbf{Rank} & \textbf{Consensus Protocol} & \textbf{Gen.} & \textbf{$P_{\text{secure}}$} & \textbf{$P_{\text{temporal}}$} & \textbf{Latency $\tau$ (ms)} & \textbf{Throughput (TPS)} & \textbf{Message Complexity $M(n)$} \\
\midrule
\textbf{1} & \textbf{Random Verifiable Rotation (RVR)} & G4 & \textbf{0.9404} & \textbf{0.0101} & \textbf{200.0} & \textbf{250.0} & $\mathcal{O}(n)$ (Linear) \\
\textbf{2} & \textbf{Tower BFT} & G4 & \textbf{0.9272} & \textbf{0.0240} & \textbf{242.9} & \textbf{205.8} & $\mathcal{O}(n)$ (Pipelined PoH) \\
3 & SV-PBFT & G3 & 0.9037 & 0.0488 & 500.0 & 100.0 & $\mathcal{O}(n^2)$ (Clustered) \\
4 & G-PBFT & G3 & 0.8902 & 0.0629 & 650.0 & 76.9 & $\mathcal{O}(n_c^2)$ (Geographic) \\
5 & CE-PBFT & G3 & 0.8770 & 0.0769 & 800.0 & 62.5 & $\mathcal{O}(n^2)$ (Elective) \\
6 & QBFT & G2 & 0.8177 & 0.1393 & 1500.0 & 33.3 & $\mathcal{O}(n^2)$ (Quorum) \\
7 & IBFT 2.0 & G2 & 0.7399 & 0.2212 & 2500.0 & 20.0 & $\mathcal{O}(n^2)$ (Istanbul) \\
8 & Classic PBFT & G1 & 0.4421 & 0.5347 & 7650.0 & 6.5 & $\mathcal{O}(n^2)$ (3-Phase) \\
9 & Oral Message $OM(m)$ & G1 & 0.0124 & 0.9869 & 43350.0 & 1.1 & $\mathcal{O}(n^{f+1})$ (Recursive Exponential) \\
\bottomrule
\end{tabularx}
\end{table*}

\subsection{Monte Carlo Verification & Reproducibility}
Across $N = 10^6$ independent Monte Carlo simulation trials (`seed = 42`), empirical sample means corroborate analytical closed-form equations to within 0.5\% relative error with 95\% confidence intervals:
\begin{itemize}[leftmargin=*]
    \item $P_{\text{SA}}$ Analytical: 0.59871 vs. Monte Carlo: 0.59888 ($\text{Error} = 0.028\%$).
    \item $P_{\text{SA, BFT}}$ Analytical: 0.35853 vs. Monte Carlo: 0.35867 ($\text{Error} = 0.040\%$).
    \item $P_{\text{MitM, BFT}}$ Analytical: 0.00250 vs. Monte Carlo: 0.00251 ($\text{Error} = 0.400\%$).
    \item $P_{\text{Key, BFT}}$ Analytical: $9.851 \times 10^{-6}$ vs. Monte Carlo: $9.848 \times 10^{-6}$ ($\text{Error} = 0.030\%$).
\end{itemize}

% ═══════════════════════════════════════════════════════════════
\section{Conclusion & Future Roadmap}
\label{sec:conclusion_restructured}
% ═══════════════════════════════════════════════════════════════

This paper presented the \emph{Static–Temporal Security Framework (STSF)}, transforming the evaluation of BFT blockchain consensus in EV energy trading networks into a single, coherent engineering story. Within baseline model assumptions, permissioned BFT consensus achieves a 170 order-of-magnitude static security gain, eliminates single-point-of-failure risks ($P_{\text{SPOF}} \to 10^{-10}$), and neutralizes replay vectors. Furthermore, sub-second pipelined consensus protocols (Tower BFT, RVR) preserve over 92\% of static security gains under real-time Poisson attack traffic.

Future research will focus on deploying our BFT framework on an experimental 5G-connected smart grid hardware-in-the-loop (HIL) testbed and evaluating dynamic validator set churn under correlated regional physical failures.

% ═══════════════════════════════════════════════════════════════
% REFERENCES
% ═══════════════════════════════════════════════════════════════
\begin{thebibliography}{99}

\bibitem{sheikh2020}
A.~Sheikh, A.~Ahmadinia, and R.~Kianoush, ``Secured energy trading using Byzantine-based blockchain consensus in smart grid,'' \emph{IEEE Access}, vol.~8, pp.~156\,832--156\,845, 2020.

\bibitem{lamport1982}
L.~Lamport, R.~Shostak, and M.~Pease, ``The Byzantine generals problem,'' \emph{ACM Trans. Program. Lang. Syst.}, vol.~4, no.~3, pp.~382--401, 1982.

\bibitem{castro1999}
M.~Castro and B.~Liskov, ``Practical Byzantine fault tolerance,'' in \emph{Proc. OSDI}, 1999, pp.~173--186.

\bibitem{li2020_blockchain}
Z.~Li, J.~Kang, R.~Yu, D.~Ye, Q.~Deng, and Y.~Zhang, ``Consortium blockchain for secure energy trading in industrial IoT,'' \emph{IEEE Trans. Ind. Inf.}, vol.~14, no.~8, pp.~3690--3700, 2018.

\bibitem{baumeister2010}
T.~Baumeister, ``Literature review on smart grid cyber security,'' Collaborative Softw. Develop. Lab. at Univ. Hawaii, Honolulu, HI, USA, Tech. Rep., 2010.

\bibitem{sridhar2012}
S.~Sridhar, A.~Hahn, and M.~Govindarasu, ``Cyber-physical system security for the electric power grid,'' \emph{Proc. IEEE}, vol.~100, no.~1, pp.~210--224, 2012.

\bibitem{wang2013}
W.~Wang and Z.~Lu, ``Cyber security in the smart grid: Survey and challenges,'' \emph{Computer Networks}, vol.~57, no.~5, pp.~1344--1371, 2013.

\bibitem{yan2012}
Y.~Yan, Y.~Qian, H.~Sharif, and D.~Tipper, ``A survey on cyber security for smart grid communications,'' \emph{IEEE Commun. Surveys Tuts.}, vol.~14, no.~4, pp.~998--1010, 2012.

\bibitem{ding2024}
X.~Ding \emph{et al.}, ``CE-PBFT: An efficient cluster-based PBFT consensus algorithm for microgrid energy trading,'' \emph{IEEE Trans. Smart Grid}, vol.~15, no.~2, pp.~1820--1832, 2024.

\end{thebibliography}

\end{document}
