\documentclass[12pt,a4paper]{article}
\usepackage{amsmath,amssymb,booktabs,graphicx,geometry,hyperref,xcolor}
\geometry{margin=1in}
\definecolor{ieee}{RGB}{0,83,159}

\title{\textbf{Week 1: Identifying Literature Gaps \& Vulnerability Discovery}\\[6pt]
\large Based on: Sheikh et al., ``Secured Energy Trading Using Byzantine-Based Blockchain Consensus,'' IEEE Access, 2020}
\author{VJTI Research Report}
\date{June 2026}

\begin{document}
\maketitle

% ══════════════════════════════════════════════════════════════════
\section{Paper Summary}
% ══════════════════════════════════════════════════════════════════

Sheikh et al.\ propose a Byzantine-based blockchain consensus mechanism for
securing energy trading between Electric Vehicles (EVs) and a Distribution
Network (DN) on an IEEE 33-bus system. The key contributions are:

\begin{itemize}
\item \textbf{System Model}: 50 EVs + 1 DN = 51 nodes, with $n_{sen} = 3854$ sensors.
\item \textbf{Energy Model} (\S IV): SOC-based participation (Eq.~3), remaining charge $C_{rem}^i$ (Eq.~5), three trading scenarios (Table~3).
\item \textbf{BFT Consensus} (\S V): Oral Message OM($m$) from Lamport et al., requiring $n \geq 3f+1$ to tolerate $f$ traitors.
\item \textbf{
<truncated 8389 bytes>
teration involves $O(n^m)$ network-wide broadcast rounds. This masks the true computational cost.

% ══════════════════════════════════════════════════════════════════
\section{Summary of Findings}
% ══════════════════════════════════════════════════════════════════

\begin{table}[htbp]
\centering
\caption{Summary of Literature Gaps and Vulnerabilities}
\begin{tabular}{@{}clp{7cm}@{}}
\toprule
\textbf{\#} & \textbf{Type} & \textbf{Description} \\
\midrule
G1 & Gap & $O(n^m)$ message complexity --- infeasible at $n{=}51$, $m{=}16$ \\
G2 & Gap & No latency analysis (43.35s consensus unquantified) \\
G3 & Gap & No TPS metrics (only 1.1 TPS) \\
G4 & Gap & No network sensitivity analysis \\
G5 & Gap & Sensor--validator architectural confusion \\
G6 & Gap & No alternative consensus comparison \\
G7 & Gap & No empirical validation (purely analytical) \\
\midrule
V1 & Vuln & 43.35s window of vulnerability for DoS/injection \\
V2 & Vuln & Recursive message amplification attack \\
V3 & Vuln & Network fragility from recursive message dependency \\
V4 & Vuln & Misleading iteration semantics \\
\bottomrule
\end{tabular}
\end{table}

\textbf{Conclusion:} These 7~gaps and 4~vulnerabilities motivate our Week~2 work:
implementing actual OM($m$) and Tower~BFT consensus simulators, validating them
with Monte Carlo trials, and comparing both algorithms across all three energy
trading scenarios using the \emph{exact same} parameters from Sheikh et al.
Our empirical approach addresses Gap~G7 directly.

\end{document}

The above content shows the entire, complete file contents of the requested file.
